---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {hydricsoils}

<!-- badges: start -->
[![R-CMD-check](https://github.com/brownag/hydricsoils/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/brownag/hydricsoils/actions/workflows/R-CMD-check.yaml)
[![hydricsoils Manual](https://img.shields.io/badge/docs-HTML-informational)](https://humus.rocks/hydricsoils/)
<!-- badges: end -->

The goal of {hydricsoils} is to provide easy access to definitions, criteria, and area of applicability for 'Field Indicators of Hydric Soils in the United States' and also to provide tools which assist in evaluating associated soil morphology.

## Installation

You can install the development version of {hydricsoils} like so:

``` r
if(!require("remotes")) install.packages("remotes")
remotes::install_github("brownag/hydricsoils")
```

## Examples

### Example 1

This basic example shows how to determine the area of applicability of indicator "A9" (also known as "1cm Muck"). 

```{r example1}
library(hydricsoils)

data(fihs, package = "hydricsoils")

subset(fihs, indicator == "A9", select = c("usage", "usage_symbols", "except_mlra"))
```

If we load the MLRA v5.2 database, for instance using {terra}, we can visualize the extent of where the "A9" indicator is used (in blue), and then show the MLRA it is excluded from (a portion of LRR P, "South Atlantic and Gulf Slope Cash Crops, Forest, and Livestock Region", in red):

```{r example1-spatial}
library(terra)
x <- vect("/vsizip//vsicurl/https://www.nrcs.usda.gov/sites/default/files/2022-10/MLRA_52_2022.zip/MLRA_52_2022")
ind <- fihs[fihs$indicator == "A9",]
xsub <- subset(x, x$LRRSYM %in% ind[, 'usage_symbols'][[1]])
plot(xsub, col = "BLUE")
plot(xsub[xsub$MLRARSYM %in% unlist(ind$except_mlra),], add = TRUE, col = "RED")
```

### Example 2

Next let's find all of the indicators that can be used in LRR "C" (also known as the "California Subtropical Fruit, Truck, and Specialty Crop Region")

```{r example2}
subset(fihs, sapply(usage_symbols, \(x) any(x == "C")), select = c("indicator", "indicator_name"))
```

This matches the information we can find in the guide in Appendix 1.

## Future work

In future updates I hope to include:

 - A defined _data.frame_ format with standard column names, data types, and relationships that are needed to evaluate criteria for all established and provisional indicators
 
 - A glossary with definitions of key criteria and terms
 
 - An index to figures and pictures from the guide, and mapping of figures to specific indicators
 
 - Parsing of LRRs and MLRAs where established indicators are being tested (new columns in existing dataset, or new dataset)
 
 - Parsing of LRRs and MLRAs where provisional indicators are being tested (new dataset)

 - Routines for automatic evaluation of input data to determine which indicators may be met
 
 - Spatial methods and helpers for creating graphics depicting where indicators are used or not
 
## Disclaimer

The data and routines in this **R** package (hereafter the "Work") are not intended as a replacement for a thorough understanding of the latest edition Field Indicators of Hydric Soils as distributed by official sources.  

This Work is released under the Creative Commons 1.0 Universal license. In particular the Work is provided "as-is" and makes no representations or warranties of any kind, express, implied, statutory or otherwise, including without limitation warranties of title, merchantability, fitness for a particular purpose, non infringement, or the absence of latent or other defects, accuracy, or the present or absence of errors, whether or not discoverable, all to the greatest extent permissible under applicable law.

## References

United States Department of Agriculture, Natural Resources Conservation Service. 2018. Field Indicators of Hydric Soils in the United States, Version 8.2. L.M. Vasilas, G.W. Hurt, and J.F. Berkowitz (eds.). USDA, NRCS, in cooperation with the National Technical Committee for Hydric Soils. Available online: <https://www.nrcs.usda.gov/resources/guides-and-instructions/field-indicators-of-hydric-soils>
