---
title: "Field Indicators of Hydric Soils in the United States"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Field Indicators of Hydric Soils in the United States}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(hydricsoils)
data(fihs, package = "hydricsoils")
show_plots <- (!inherits(try(requireNamespace("terra", quietly = TRUE), silent = TRUE), 'try-error')  && file.exists(file.path(tools::R_user_dir("hydricsoils", "data"), "lrrmlra.gpkg")))

if (show_plots) {
  all_ext <- lrrmlra_geometry()
  conus_ext <- terra::project(terra::subset(all_ext, !all_ext$LRRSYM %in% c("V", "Q", "W1", "W2", "X1", "X2", "Y", "Z")), "EPSG:5070")
  ak_ext <- terra::project(terra::subset(all_ext, all_ext$LRRSYM %in% c("W1", "W2", "X1", "X2", "Y")), "EPSG:3338")
}
```

This vignette shows the usage, criteria, and notes related to the application of Field Indicators of Hydric Soils in the United States.

Usage ("**Used in:**" section) and testing ("**Test in:**" section) of hydric indicators are generally based on Land Resource Region (denoted by a code with a single letter, or a single letter followed by a single digit). In some cases indicators are used only in specific MLRAs (denoted by a code starting with one or more digits, possibly ending with a single letter) of a region. In other cases, an indicator is used in all parts of a Land Resource Region except for specific MLRAs ("**Except MLRA:**" section)

Extent maps for indicators that are _not_ specific to Alaska are limited to CONUS (using coordinate reference system `"EPSG:5070"`) for clarity. Alaskan indicators are limited to the Alaskan Land Resource Regions (using `"EPSG:3338"`). See `plot_indicator_usage()` to create maps using the full extent (i.e. for details on applicability in Pacific Islands or the Caribbean).

```{r fihs1, echo=FALSE, results='asis', fig.width=8, fig.height=4, fig.align='center', out.width="100%"}
for (i in seq(nrow(fihs))) {
  f <- fihs[i, ]
  cat("##", f$indicator, " -- ", f$indicator_name, "\n\n")
  
  cat(f$usage, "\n\n")
  cat(
    "**Used in:** ",
    paste0(f$usage_symbols[[1]], collapse = ", "),
    ifelse(
      length(f$except_mlra[[1]]),
      paste0("\n\n - **Except MLRA:** ", paste0(f$except_mlra[[1]], collapse = ", ")),
      ""
    ),
    "\n\n"
  )
  if (length(f$test_symbols[[1]]) || length(f$test_except_mlra[[1]])) {
    cat(
      "**Test in:** ",
      ifelse(length(f$test_symbols[[1]]),
             paste0(f$test_symbols[[1]], collapse = ", "),
             ""), 
      ifelse(
        length(f$test_except_mlra[[1]]),
        paste0("\n\n - **Except MLRA:** ", paste0(f$test_except_mlra[[1]], collapse = ", ")),
        ""
      ),
      "\n\n"
    )
  }
  cat("**Criteria:**\n", f$criteria)
  if (length(f$notes))
    cat("**User Notes:**\n", f$notes)
  cat("\n\n")
  
  if (show_plots) {
     if (grepl("Alaska", f$indicator_name)) {
       try(plot_indicator_usage(f$indicator, test_areas = TRUE, ext = ak_ext, crs = "EPSG:3338"), silent = FALSE)
     } else {
       try(plot_indicator_usage(f$indicator, test_areas = TRUE, ext = conus_ext, crs = "EPSG:5070"), silent = FALSE)
     }
     
  }
  cat("\n\n")
  
}
```
